Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-09-19 paper/paper.tex
edited Mannik paragrph in Introduction Tested everything on quipu.

Previous studies have shown that the minD protein system is capable of
exhibiting oscillations in round shapes \cite{fange2006noise}as well
as in connected three pronged tube shapes~\cite{varma2008min}, in
which the oscillations seem to seek out the extreme poles in the
cell~\cite{corbin2002exploring} \cite{juarez2010changes}. However,
Mannik \emph{et al.} have also shown that there are limitations to
this capability~\cite{mannik2010bacteria}
\cite{mannik2009bacterial}. They have experimentally forced E. Coli
cells into flattened irregular shapes of $0.25\mu m$
thickness. Oscillations in this type of cell shape lose their
regularity, both temporally and spatially. This allows for an
opportunity to test previously studied MinD simulation models against
more extreme experimental cases than have been seen thus
far. \fixme{Perhaps this sentence should come later after we've
  introduced the simulation research that's been done.}

>>>>

Previous studies have shown that the minD protein system is capable of
exhibiting oscillations in round shapes \cite{fange2006noise}as well
as in connected three pronged tube shapes~\cite{varma2008min}, in
which the oscillations seem to seek out the extreme poles in the
cell~\cite{corbin2002exploring,juarez2010changes}.

However, Mannik \emph{et al.} have also shown that there are
limitations to this robust capability to
oscillate~\cite{mannik2010bacteria,mannik2009bacterial}. They have
experimentally forced E. Coli cells into silicon microfabricated
channels of $0.25\mu m$ thickness. Upon entering the channels, the
cells undergo a rapid mechanical deformation in which they widen along
their short axis parallel to the channel plane, followed by a slower,
broadening out along both free axes.  This deformation results in very
wide (they can reach widths of over $5\mu m$), flattened cells that
when view from the top down show very irregular and assymetric shapes.
While these cells are still able to divide into surprisingly equal
volumes, the MinD oscillations within lose their regularity, both
temporally and spatially. Seen with flourescent microscopy (after
using common genetic techniques to allow the cells to flouresce), the
MinD seem to maximize in multiple locations within the cell, in a
seemingly random sequence. This allows for an opportunity to test
previously studied MinD simulation models against more extreme
experimental cases than have been seen thus far. \fixme{Perhaps this
  sentence should come later after we've introduced the simulation
  research that's been done.}
